The not so widening achievement gap: an interntational comparison of the evolution of the SES achievement gap from 2000-2015

### 2017-08-18 paper/paper.Rnw
Added 80/20 and 70/30 graphs at the end - still need to check if they work

% Add confidence intervals
\begin{figure}
\begin{center}
<<country_achievement_and_inequality, include = T, out.height = '5in', out.width = '5.5in', fig.align = 'center'>>=
data_to_plot %>%
  mutate(Mean = scale(Mean),
         country = factor(country, levels = ordered_cnt, ordered = TRUE)) %>%
  gather(type, value, difference, Mean) %>%
  filter(country %in% countries) %>%
  ggplot(aes(as.character(wave), value, group = type, colour = type, shape = type)) +
  geom_point(size = 2, alpha = 0.4) +
  geom_line(alpha = 0.4) +
  scale_y_continuous(name = "Test scores in SD") +
  scale_x_discrete(name = NULL, breaks = c(2000, 2009, 2015)) +
  scale_colour_discrete(name = NULL, labels = c("Achievement gap", "Avg performance")) +
  scale_shape_discrete(name = NULL, labels = c("Achievement gap", "Avg performance")) +
  ggtitle("Evolution of the top/bottom groups by different gaps") +
  facet_wrap(~ country, ncol = 5) +
  theme_few() +
  theme(panel.spacing = unit(1, "lines"),
        legend.position = "bottom")
@
\caption{Time trends of achievement gap and average achiement between 2000 and 2015}
\end{center}
\end{figure}

@@

<<>>=
results_math <- test_diff(adapted_year_data, reliability_pisa, "MATH", c(0.1, 0.9))
results_read <- test_diff(adapted_year_data, reliability_pisa, "READ", c(0.1, 0.9))
results_math_topmid <- test_diff(adapted_year_data, reliability_pisa, "MATH", c(0.5, 0.9))
results_read_topmid <- test_diff(adapted_year_data, reliability_pisa, "READ", c(0.5, 0.9))
results_math_midbottom <- test_diff(adapted_year_data, reliability_pisa, "MATH", c(0.1, 0.5))
results_read_midbottom <- test_diff(adapted_year_data, reliability_pisa, "READ", c(0.1, 0.5))

@@

>>>>

% Add confidence intervals
\begin{figure}
\begin{center}
<<country_achievement_and_inequality, include = T, out.height = '5in', out.width = '5.5in', fig.align = 'center'>>=
data_to_plot %>%
  mutate(Mean = scale(Mean),
         country = factor(country, levels = ordered_cnt, ordered = TRUE)) %>%
  gather(type, value, difference, Mean) %>%
  filter(country %in% countries) %>%
  ggplot(aes(as.character(wave), value, group = type, colour = type, shape = type)) +
  geom_point(size = 2, alpha = 0.4) +
  geom_line(alpha = 0.4) +
  scale_y_continuous(name = "Test scores in SD") +
  scale_x_discrete(name = NULL, breaks = c(2000, 2009, 2015)) +
  scale_colour_discrete(name = NULL, labels = c("Achievement gap", "Avg performance")) +
  scale_shape_discrete(name = NULL, labels = c("Achievement gap", "Avg performance")) +
  ggtitle("Evolution of the top/bottom groups by different gaps") +
  facet_wrap(~ country, ncol = 5) +
  theme_few() +
  theme(panel.spacing = unit(1, "lines"),
        legend.position = "bottom")
@
\caption{Time trends of achievement gap and average achievement between 2000 and 2015}
\end{center}
\end{figure}

@@

<<estimating_othergaps>>=
results_math_80 <- test_diff(adapted_year_data, reliability_pisa, "MATH", c(0.8, 0.2))
results_read_80 <- test_diff(adapted_year_data, reliability_pisa, "READ", c(0.8, 0.2))
results_math_topmid_80 <- test_diff(adapted_year_data, reliability_pisa, "MATH", c(0.8, 0.5))
results_read_topmid_80 <- test_diff(adapted_year_data, reliability_pisa, "READ", c(0.8, 0.5))
results_math_midbottom_80 <- test_diff(adapted_year_data, reliability_pisa, "MATH", c(0.5, 0.2))
results_read_midbottom_80 <- test_diff(adapted_year_data, reliability_pisa, "READ", c(0.5, 0.2))

results_math_70 <- test_diff(adapted_year_data, reliability_pisa, "MATH", c(0.70, 0.30))
results_read_70 <- test_diff(adapted_year_data, reliability_pisa, "READ", c(0.70, 0.30))
results_math_topmid_70 <- test_diff(adapted_year_data, reliability_pisa, "MATH", c(0.70, 0.5))
results_read_topmid_70 <- test_diff(adapted_year_data, reliability_pisa, "READ", c(0.70, 0.5))
results_math_midbottom_70 <- test_diff(adapted_year_data, reliability_pisa, "MATH", c(0.5, 0.30))
results_read_midbottom_70 <- test_diff(adapted_year_data, reliability_pisa, "READ", c(0.5, 0.30))

@@

\begin{figure}
\begin{center}
<<graph_all_gaps>>=
complete_data_topbottom_80 <-
  pisa_preparer(results_math_80, results_read_80) %>%
  mutate(type = "80th/20th SES gap")

complete_data_topmid_80 <-
  pisa_preparer(results_math_topmid_80, results_read_topmid_80) %>%
  mutate(type = "80th/50th SES gap")

complete_data_midbottom_80 <-
  pisa_preparer(results_math_midbottom_80, results_read_midbottom_80) %>%
  mutate(type = "50th/20th SES gap")

######

complete_data_topbottom_70 <-
  pisa_preparer(results_math_70, results_read_70) %>%
  mutate(type = "70th/30th SES gap")

complete_data_topmid_70 <-
  pisa_preparer(results_math_topmid_70, results_read_topmid_70) %>%
  mutate(type = "70th/50th SES gap")

complete_data_midbottom_70 <-
  pisa_preparer(results_math_midbottom_70, results_read_midbottom_70) %>%
  mutate(type = "50th/30th SES gap")

#####

complete_gaps <-
  complete_data_topbottom %>%
  bind_rows(complete_data_topbottom_80)
  bind_rows(complete_data_topbottom_70)

complete_gaps %>%
  mutate(country = factor(country, levels = ordered_cnt, ordered = TRUE)) %>%
  filter(type_test == "math",
         country %in% countries) %>%
  ggplot(aes(as.character(wave), difference, colour = type, linetype = type, group = type)) +
  geom_line(stat = "smooth", method = "lm", size = 0.7) +
  facet_wrap(~ country, ncol = 5) +
  scale_colour_discrete(name = "Type of gaps") +
  scale_linetype_discrete(name = "Type of gaps") +
  scale_y_continuous(name = "Gaps in SD", expand = c(0, 0), lim = c(0, 3)) +
  scale_x_discrete(name = NULL, breaks = c(2000, 2009, 2015)) +
  ggthemes::theme_few() +
  theme(panel.spacing = unit(1, "lines"),
        panel.grid.major.y = element_line(colour = "grey"),
        legend.title = ) +
  ggtitle("Evolution of the different achievement gaps")
@
\caption{Evolution of the achievement gap for several gaps}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
<<percentage_change_8020>>=
perc_increase_fun(complete_data_topbottom_80) %>%
  perc_graph("math",
             "80/20 achievement gap",
             "Percentage change from 2000 to 2015")
@
\caption{Percentage change in the 80/20 achievement gap from 2000 to 2015}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
<<percentage_change_7030>>=
perc_increase_fun(complete_data_topbottom_70) %>%
  perc_graph("math",
             "70/30 achievement gap",
             "Percentage change from 2000 to 2015")
@
\caption{Percentage change in the 70/30 achievement gap from 2000 to 2015}
\end{center}
\end{figure}
